Reanalysis Step 3: Propensity Score Analysis with Twang

### 2025-08-17 reanalysis-3-propensity_scores.qmd
es mean only

# Run twang for each imputed dataset
message("\nCalculating propensity scores using Twang GBM...")
message("Parameters:")
message("  - n.trees: ", n_trees)
message("  - interaction.depth: 3")
message("  - shrinkage: 0.01")
message("  - estimand: ATT")
message("  - stop.method: es.mean, ks.max")

@@

  # Run twang ps model
  twang_model <- ps(
    formula = ps_formula,
    data = imputed_datasets[[i]],
    n.trees = n_trees,
    interaction.depth = 3,
    shrinkage = 0.01,
    estimand = "ATT",
    stop.method = c("es.mean", "ks.max"),
    n.minobsinnode = 10,
    verbose = FALSE
  )

  # Extract weights for ATT (using es.mean as primary stop method)
  weights <- get.weights(twang_model, stop.method = "es.mean")

  # Store results
  twang_results[[i]] <- list(
    model = twang_model,
    weights = weights,
    ps_scores = twang_model$ps[, "es.mean"],
    n_trees_used = twang_model$desc$es.mean$n.trees,
    balance_es_mean = twang_model$balance$es.mean,
    balance_ks_max = twang_model$balance$ks.max
  )

  message("  - Trees used (es.mean): ", twang_model$desc$es.mean$n.trees)
  message("  - Trees used (ks.max): ", twang_model$desc$ks.max$n.trees)
}

@@

for (i in 1:n_imputations) {
  # Get balance table from twang
  bal_es <- twang_results[[i]]$balance_es_mean
  bal_ks <- twang_results[[i]]$balance_ks_max

  # Summary statistics for es.mean method
  balance_summaries[[i]] <- data.frame(
    Imputation = i,
    Mean_ES = mean(abs(bal_es$std.eff.sz), na.rm = TRUE),
    Max_ES = max(abs(bal_es$std.eff.sz), na.rm = TRUE),
    N_Imbalanced = sum(abs(bal_es$std.eff.sz) > 0.1, na.rm = TRUE),
    Mean_KS = mean(bal_ks$ks, na.rm = TRUE),
    Max_KS = max(bal_ks$ks, na.rm = TRUE)
  )
}

@@

```{r summary}
# Final summary
message("\n=== TWANG PROPENSITY SCORE ANALYSIS SUMMARY ===")
message("Method: Generalized Boosted Models (GBM) via twang")
message("Imputations processed: ", n_imputations)
message("Analysis mode: ", analysis_mode)
message("Trees used: ", n_trees)
message("Estimand: ATT (Average Treatment Effect on the Treated)")
message("Stop methods: es.mean (effect size), ks.max (Kolmogorov-Smirnov)")

>>>>

# Run twang for each imputed dataset
message("\nCalculating propensity scores using Twang GBM...")
message("Parameters:")
message("  - n.trees: ", n_trees)
message("  - interaction.depth: 3")
message("  - shrinkage: 0.01")
message("  - estimand: ATT")
message("  - stop.method: es.mean (effect size only)")

@@

  # Run twang ps model with es.mean only
  twang_model <- ps(
    formula = ps_formula,
    data = imputed_datasets[[i]],
    n.trees = n_trees,
    interaction.depth = 3,
    shrinkage = 0.01,
    estimand = "ATT",
    stop.method = "es.mean",  # Only use es.mean for faster computation
    n.minobsinnode = 10,
    verbose = FALSE
  )

  # Extract weights for ATT
  weights <- get.weights(twang_model, stop.method = "es.mean")

  # Store results
  twang_results[[i]] <- list(
    model = twang_model,
    weights = weights,
    ps_scores = twang_model$ps[, "es.mean"],
    n_trees_used = twang_model$desc$es.mean$n.trees,
    balance_es_mean = twang_model$balance$es.mean
  )

  message("  - Trees used (es.mean): ", twang_model$desc$es.mean$n.trees)
}

@@

for (i in 1:n_imputations) {
  # Get balance table from twang (es.mean only)
  bal_es <- twang_results[[i]]$balance_es_mean

  # Summary statistics for es.mean method
  balance_summaries[[i]] <- data.frame(
    Imputation = i,
    Mean_ES = mean(abs(bal_es$std.eff.sz), na.rm = TRUE),
    Max_ES = max(abs(bal_es$std.eff.sz), na.rm = TRUE),
    N_Imbalanced = sum(abs(bal_es$std.eff.sz) > 0.1, na.rm = TRUE),
    # KS statistics are still available in the es.mean balance table
    Mean_KS = mean(bal_es$ks, na.rm = TRUE),
    Max_KS = max(bal_es$ks, na.rm = TRUE)
  )
}

@@

```{r summary}
# Final summary
message("\n=== TWANG PROPENSITY SCORE ANALYSIS SUMMARY ===")
message("Method: Generalized Boosted Models (GBM) via twang")
message("Imputations processed: ", n_imputations)
message("Analysis mode: ", analysis_mode)
message("Trees used: ", n_trees)
message("Estimand: ATT (Average Treatment Effect on the Treated)")
message("Stop method: es.mean (effect size minimization)")
